Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-12-18 MDPI_submission.Rmd
Commit 2023-12-18 19:10

The linear trends were calculated using an autoregressive model
<!-- ([ARIMA......)]{.mark} and were reported together with the 2σ errors. -->
<!-- [For the trends derived from monthly mean data we used an effective -->
<!-- number of observations to calculate the standard error (Sander et al., -->
<!-- 2000) in order to assess the significance of the trends.]{.mark} -->
@Santer2000

@@

## numeric precision of table
LT$slope    <- signif(LT$slope,   4)
LT$slope.p  <- signif(LT$slope.p, 4)
LT$Rsqrd    <- signif(LT$Rsqrd,   4)

@@

LT <- rename(LT, "Trend [%/year]" = slope       )
LT <- rename(LT, "Sky conditions" = DATA        )
LT <- rename(LT, "Days"           = N           )
LT <- rename(LT, "Pearson cor."   = cor.estimate)
LT <- rename(LT, "Trend p-value"  = slope.p     )

LT$slope.stat_sig  <- NULL
LT$`Trend p-value` <- NULL
LT$N_eff           <- NULL

# LT$`Trend statistical signif. [%]` <- NULL

## order columns
# LT1 <- LT[,c(
#       "Sky conditions",
#       "Trend [%/year]",
#       "Trend p-value",
#       "Pearson cor.",
#       "Days"
#       )]
LT1 <- LT

@@

x_latex <- knitr::kable(LT1,
                        format      = "latex",
                        booktabs    = TRUE,
                        escape      = TRUE,
                        format.args = list(digits = 3),
                        linesep     = "",
                        align       = "crrrr",
                        centering   = FALSE,
                        table.envir = "table",
                        position    = "H",
                        caption     = paste0("Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))

@@

For cloudy-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA as the variability of irradiance is
dominated by the cloud effects.
For SZAs $25^\circ$ - $75^\circ$, the trends range within about $\pm 0.2\,\%/y$, 
the trends are variable and for most SZAs insignificant ranging between
$0$ and $0.2%\,\%/y$.
Between $50^\circ$ and $75^\circ$ SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.

>>>>

The linear trends were calculated using an first order autoregressive model
(arima) with lag-1, and were reported together with the $2\sigma$ errors (Table\ \ref{tab:trendtable}).

<!-- [For the trends derived from monthly mean data we used an effective -->
<!-- number of observations to calculate the standard error (Sander et al., -->
<!-- 2000) in order to assess the significance of the trends.]{.mark} -->
@Santer2000

@@

## numeric precision of table
LT$slope         <- signif(LT$slope,         3)
LT$slope.p       <- signif(LT$slope.p,       3)
LT$Rsqrd         <- signif(LT$Rsqrd,         3)
LT$Tmod_Estimate <- signif(LT$Tmod_Estimate, 3)
LT$cor.estimate  <- signif(LT$cor.estimate,  3)
LT$Tmod_Std..Error <- signif(LT$Tmod_Std..Error * 2, 3)
LT$ChangeWpY       <- signif(LT$ChangeWpY, 3)

@@

# LT <- rename(LT, "Trend [%/year]" = slope        )
LT <- rename(LT, "Sky conditions"  = DATA         )
LT <- rename(LT, "Days"            = N            )
LT <- rename(LT, "Pearson cor."    = cor.estimate )
LT <- rename(LT, "Trend p-value"   = slope.p      )
LT <- rename(LT, "Trend [%/year]"  = Tmod_Estimate) # Arima slope
LT <- rename(LT, "Trend S.E."      = Tmod_Std..Error )
LT <- rename(LT, "Trend [W/year]"  = ChangeWpY )

LT$slope.stat_sig  <- NULL
LT$`Trend p-value` <- NULL
LT$N_eff           <- NULL
LT$slope           <- NULL
LT$t_eff           <- NULL
LT$t_eff_cri       <- NULL
LT$conf_2.5        <- NULL
LT$conf_97.5       <- NULL
LT$mean_clima      <- NULL
LT$Tmod_z.value    <- NULL
LT$Tmod_Pr...z..   <- NULL

## order columns
LT1 <- LT[,c(
      "Sky conditions",
      "Trend [%/year]",
      "Trend S.E.",
      "Pearson cor.",
      "Trend [W/year]",
      "Days"
      )]
# LT1 <- LT

@@

# names(LT1)
# "Sky conditions", "Trend [%/year]", "Trend S.E.", "Pearson cor.", "Trend [W/year]", "Days" 
x_latex <- knitr::kable(LT1,
                        format      = "latex",
                        booktabs    = TRUE,
                        escape      = FALSE,
                        col.names   = c("Sky conditions", "Trend [\\%/year]", "Trend S.E. ($2\\sigma$)", "Pearson cor.", "Trend [W/year]", "Days"),
                        format.args = list(digits = 3),
                        linesep     = "",
                        align       = "crrrrr",
                        centering   = FALSE,
                        table.envir = "table",
                        position    = "H",
                        caption     = paste0("Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))

@@

For cloudy-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA as the variability of irradiance is
dominated by the cloud effects.
For SZAs $25^\circ$ -- $75^\circ$, the trends range within about $\pm 0.2\,\%/y$,
the trends are variable and for most SZAs insignificant ranging between $0$ and $0.2\,\%/y$.
The smooth increase of the trend at SZAs larger than $75^\circ$ is probably associated with stronger attenuation by
clouds under oblique incidence angles which results in smaller
variability.

### 2023-12-18 MDPI_submission.tex
Commit 2023-12-18 19:10

The linear trends were calculated using an autoregressive model
\citet{Santer2000}

@@

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Trend [\%/year] & Pearson cor. & Sky conditions & Days & t\_eff & t\_eff\_cri & conf\_2.5 & conf\_97.5 & mean\_clima & ChangeWpY\\
\midrule
0.3756 & 0.0911 & All skies & 10250 & 4.727 & 1.96 & 0.000811 & 0.001246 & 387 & 1.452\\
0.0972 & 0.1383 & Clear skies & 2684 & 0.476 & 1.96 & 0.000194 & 0.000338 & 515 & 0.501\\
0.4077 & 0.0810 & Cloudy skies & 4937 & 3.566 & 1.96 & 0.000733 & 0.001499 & 290 & 1.181\\
\bottomrule
\end{tabu}
\end{table}

@@

For cloudy-sky conditions (Figure~\ref{fig:szatrends-3}), we can not
discern any significant dependence of the SDR trend with SZA as the
variability of irradiance is dominated by the cloud effects. For SZAs
\(25^\circ\) - \(75^\circ\), the trends range within about
\(\pm 0.2\,\%/y\), the trends are variable and for most SZAs
insignificant ranging between \(0\) and \(0.2%\,\%/y\). Between
\(50^\circ\) and \(75^\circ\) SZA the trends for the period before noon
are stronger and negative, possibly associated with stronger attenuation
by clouds under oblique incidence angles.

@@

\begin{verbatim}
        <!-- % \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2}  % Seasonal from Daily -->
\end{verbatim}

>>>>

The linear trends were calculated using an first order autoregressive
model (arima) with lag-1, and were reported together with the
\(2\sigma\) errors (Table~\ref{tab:trendtable}).

\citet{Santer2000}

@@

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson cor. & Trend [W/year] & Days\\
\midrule
All skies & 0.3760 & 0.1210 & 0.0911 & 1.450 & 10250\\
Clear skies & 0.0972 & 0.0327 & 0.1380 & 0.501 & 2684\\
Cloudy skies & 0.4080 & 0.1770 & 0.0810 & 1.180 & 4937\\
\bottomrule
\end{tabu}
\end{table}

@@

For cloudy-sky conditions (Figure~\ref{fig:szatrends-3}), we can not
discern any significant dependence of the SDR trend with SZA as the
variability of irradiance is dominated by the cloud effects. For SZAs
\(25^\circ\) -- \(75^\circ\), the trends range within about
\(\pm 0.2\,\%/y\), the trends are variable and for most SZAs
insignificant ranging between \(0\) and \(0.2\,\%/y\). The smooth
increase of the trend at SZAs larger than \(75^\circ\) is probably
associated with stronger attenuation by clouds under oblique incidence
angles which results in smaller variability.

@@
